---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# addressr

<!-- badges: start -->
<!-- badges: end -->

The goal of addressr is to standardize address cleaning for various datasets used by the Eviction Lab.

## Installation

You can install the development version of addressr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("EvictionLab/addressr")
```

## Usage

Addresses can be difficult to work with and messy in many different ways. 

Here is an example of a relatively clean table:

```{r dfExample}
address_table <- dplyr::tribble(
  ~address, ~city, ~state,
  "456 Jersey Avenue #102", "Montclair", "NJ",
  "123-125 N Street Rd", "Cincinnati", "OH",
  "789 Pirate Cv East", "Memphis", "TN"
)
```

Geocoders can be thrown off by various things, such as address ranges, unit numbers, directionals, and street endings.

One method of standardizing addresses is to piece apart an address column and put it back together in friendlier format for geocoders.

The `extract_remove_squish()` allows you to tweeze out parts of the address using regex. The pattern will be removed from one column, placed into a new column, and the original column will be trimmed of whitespace. Trimming whitespace ensures we can chain certain regex patterns, notably `^` or `$` to indicate the start or end of the string.

To use the function, call `extract_remove_squish(original_column, new_column, pattern)`. 

```{r, message=FALSE}
library(dplyr)
library(addressr)

address_table <- address_table |>
  extract_remove_squish(address, street_number_range, "^\\d+-\\d+\\b") |>
  extract_remove_squish(address, street_number, "^\\d+\\b") |>
  extract_remove_squish(address, unit, "#.*") |>
  select(street_number, street_number_range, street_name = address, unit, everything())

address_table
```

This method is flexible and allows you to:

- Adjust the order of extractions
- Edit the regex easily, or declare the regex separately and call it within the function.

### `address_abbreviations` dataset

One common issue with address cleaning is abbreviation of directions, street names, and street endings. The `address_abbreviations` dataset assists by providing a resource of common abbreviations.

```{r}
slice_head(address_abbreviations, n = 3, by = type)
```

The table has three columns: `short` contains the abbreviation, `long` contains the full word, and `type` is the category of the word/abbreviation. There are currently four categories:

- `directions` for N, S, E, W, NE, NW, SE, and SW
- `numbered_street_name` for ordinal street names like 1st, 2nd, or 3rd St.
- `all_street_suffix` and `official_street_suffix` are derived primarily from the [USPS's list of officially accepted street endings](https://pe.usps.com/text/pub28/28apc_002.htm). Both contain the same abbreviated values. For the code behind this chunk, see `data-raw/address_abbreviations.R`.
  - `all_street_suffix` contains all accepted spellings of the street suffix (one-to-many). Good for catching varied spellings of street endings and reducing them to a common abbreviation.
  - `official_street_suffix` contains only the official spellings of the un-abbreviated street suffix (one-to-one). Good for expanding abbreviated street endings to a common spelling.
  
Planned future additions: Commonly abbreviated street names (MLK Jr Blvd).

### `str_collapse_bound()` to create a regex string

Extract directions from an address by using the values in `address_abbreviations`. To use `str_extract()` or `extract_remove_squish()`, all directionals need to be flattened into one string. 

The `str_collapse_bound()` function:

- Flattens strings, separating each with an OR (`|`) operator. ex: `N|S|E|W`
- Places word boundaries around the strings. Adding the boundaries ensures random letters are not extracted from the middle of words. ex: `\\b(N|S|E|W)\\b`

To catch both short and long spellings of directions, include both columns in the function.

```{r}
directions <- address_abbreviations |> filter(type == "directions")
directions <- str_collapse_bound(c(directions$short, directions$long))

# values in address_abbreviations are exclusively upper case
address_table <- address_table |> mutate(street_name = toupper(street_name))

address_table |> 
  extract_remove_squish(street_name, direction, directions) |> 
  select(direction, street_name)
```

The function worked, but pre-directions and post-directionals should ideally be in separate columns. We can edit the regex pattern to specify which direction to extract.

```{r}
regex_pre_direction <- paste0("^", directions)
regex_post_direction <- paste0(directions, "$")

address_table <- address_table |> 
  extract_remove_squish(street_name, pre_direction, regex_pre_direction) |> 
  extract_remove_squish(street_name, post_direction, regex_post_direction)

address_table |> select(pre_direction, street_name, post_direction)
```

### Alter strings with `switch_abbreviations()` and `str_replace_names()`

The `switch_abbreviation()` function can be used to abbreviate or un-abbreviate strings. By default, the function abbreviates strings.

```{r}
address_table <- address_table |> 
  mutate(post_direction = switch_abbreviation(post_direction, "directions"))

address_table |> select(pre_direction, street_name, post_direction)
```

Street endings can also be extracted using the `address_abbreviations` dataset. To expand abbreviations, change `method = "short-to-long"`.

```{r}
street_suffixes <- address_abbreviations |> filter(type == "all_street_suffix")
street_suffixes <- paste0("(?<!^)", str_collapse_bound(unique(c(street_suffixes$short, street_suffixes$long))), "$")

address_table <- address_table |> 
  extract_remove_squish(street_name, street_suffix, street_suffixes) |> 
  mutate(street_suffix = switch_abbreviation(street_suffix, "all_street_suffix", "short-to-long"))

address_table |> select(street_name, street_suffix)
```

The `switch_abbreviation()` function was made to be used with the `address_abbreviations` table. To use your own input and output, use `str_replace_names(string, input, output)`.

```{r}
address_table |> 
  mutate(state_long = str_replace_names(state, state.abb, state.name)) |>
  select(city, state_short = state, state_long)
```

More to come.
