---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# addressr

<!-- badges: start -->
<!-- badges: end -->

The goal of addressr is to standardize address cleaning for various datasets used by the Eviction Lab.

## Installation

You can install the development version of addressr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("EvictionLab/addressr")
```

## Usage

Addresses can be difficult to work with and messy in many different ways. 

Here is an example of a relatively clean table:

```{r dfExample}
address_table <- dplyr::tribble(
  ~address, ~city, ~state,
  "456 Jersey Avenue #102", "Montclair", "NJ",
  "123-125 N Street Rd", "Cincinnati", "OH",
  "789 Pirate Cv East", "Memphis", "TN",
  "3548 1ST ST FL 1", "St. Louis", "MO"
)
```

Geocoders can be thrown off by various things, such as address ranges, unit numbers, directionals, and street endings.

The `clean_address()` function streamlines address cleaning in a standardized pattern:

- Convert all characters to uppercase
- Tweeze apart the address column into components
- Rename ordinal street names ("1ST" = "FIRST")

```{r message=FALSE}
library(addressr)
library(dplyr)

address_out <- address_table[, 1] |> 
  clean_address(address) |> 
  # clean_address(address, dataset = "default_db") |>   # for duckdb tbls
  select(street_number, street_number_range, pre_direction, 
         street_name = address, all_street_suffix, post_direction, unit)

address_out
```


Behind the function is `extract_remove_squish()`, which will 

- input `.data` with a column of address data
- extract a regex pattern from the `original_column`
- output the extracted string into a `new_column`
- trim the original column of extra whitespace or non-word characters (`\\W+`)
- the `pattern` may be a string or a named address component, from one of the tables mentioned above

To use the function, call `extract_remove_squish(original_column, new_column, pattern)`. 

```{r, message=FALSE}
address_table |>
  extract_remove_squish(address, unit, "unit") |>
  # extract_remove_squish_db("original_column", "new_column", "pattern") # duckdb
  select(address, unit, everything())
```

Address components are defined in `data-raw`:

- `address_abbreviations` are common groups of address-related words, including official USPS street endings. 
  - Note: The abbreviations table calls `str_collapse_bound()`, to flatten values into a bounded regex,
  e.g. `\\b(N|S|E|W|NORTH|SOUTH...)\\b`. 
  Vectors call `str_collapse_bound()` by default; regex and character strings will not.
- `address_regex` defines address components, e.g. `^\\d+\\b` for `street_number`.
- `address_corrections` is available but not currently in use.

Combine `switch_abbreviation()` with the `address_abbreviations` table to 
alternate between short and long spellings of words.
By default, the function abbreviates words (`"long-to-short"`).

```{r}
address_out |> 
  # switch_abbreviation_db(column, type, method = "long-to-short") |> 
  mutate(post_direction = switch_abbreviation(post_direction, "directions")) |> 
  mutate(all_street_suffix = switch_abbreviation(all_street_suffix, "official_street_suffix", "short-to-long"))
```

To use your own input and output, use `str_replace_names(string, input, output)`.

```{r}
address_table |> 
  mutate(state_name = str_replace_names(state, state.abb, state.name)) |>
  select(city, state, state_name)
```

More to come. Please report issues, suggest/commit changes, open a discussion.
